### 2021-11-22 paper.tex
Offsetting definitions

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{todonotes}
\usepackage{enumerate}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{algpseudocode}
\usepackage{algorithm}

\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}

\newcommand{\stateSpace}{\mathcal{S}}
\newcommand{\actionSpace}{\mathcal{A}}
\newcommand{\rewardSpace}{\mathcal{R}}
\newcommand{\trasitionFunction}{T}
\newcommand{\rewardFunction}{R}
\newcommand{\modifiedRewardFunction}{\rewardFunction_{\mathrm{SEA}}}
\newcommand{\goalRewardFunction}{\rewardFunction_{\mathrm{Goal}}}
\newcommand{\generalRewardFunction}{\rewardFunction_{\mathrm{General}}}
\newcommand{\corruption}{C}
\newcommand{\deviationFromBaseline}{d}
\newcommand{\baseline}{b}
\newcommand{\tuple}[1]{\left\langle #1 \right\rangle}
\newcommand{\rawMDP}{\stateSpace, \actionSpace, \rewardSpace, \trasitionFunction, \rewardFunction}
\newcommand{\MDP}{\tuple{\rawMDP}}
\newcommand{\rawCRMDP}{\rawMDP, \corruption}
\newcommand{\CRMDP}{\tuple{\rawCRMDP}}
\newcommand{\rawSCRMDP}{\rawCRMDP, d, \LV}
\newcommand{\SCRMDP}{\tuple{\rawSCRMDP}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\absoluteValue}[1]{\left\lvert #1 \right\rvert}

\newtheorem{example}{Example}

\title{Impact offsetting}
\author{Us}
\date{2021}

\begin{document}

\maketitle

\section{Setting}

 To define offsetting lets first formalize the notion of a misspecified reward. I'm using the below formulation because I already had it lying around.

	\begin{definition}[CRMDP]
		A \emph{Corrupt Reward Markov Decision Process} (CRMDP) is a tuple $\CRMDP$, where $\stateSpace$ is the set of states,
		$\actionSpace$ is the set of actions, $\rewardSpace$ is the set of rewards, $\trasitionFunction \colon \stateSpace \times \actionSpace \to P\left( \stateSpace \right)$
		is the transition function, $\rewardFunction \colon \stateSpace \to \rewardSpace$ is the reward function and $\corruption \colon \stateSpace \to \rewardSpace$ is the corrupt
		reward. We call $\MDP$ the \emph{underlying MDP} of a given CRMDP.
		\label{def:CRMDP}
	\end{definition}

	To model having imprecise rewards we assume we only have access to $\corruption$, while we actually care about the real reward, which we are not able to specify.

	One approach to mitigating the problems arising from the discrepancy is side effect avoidance. We assume that the corrupt reward somewhat correctly specifies the actual goal,
	but the negative effects of changing the environment are underspecified. Then problematic behaviour might be reduced by correcting for avoiding any side effects of the agent's
	actions, compared to some baseline. This, however, introduces other biases, namely towards the baseline. We call behaviour stemming from these biases ``offsetting''.

	\begin{definition}[Side effect avoidance]
		\emph{Side effect avoidance} is a class of approaches for mitigating reward misspecification, which
		modify the reward passed to the agent according to the equation
		\begin{equation}
			\modifiedRewardFunction\left( \left( s_i \right)_{i=0}^t \right) =
			\corruption\left( s_t \right) - 
			\lambda \cdot \deviationFromBaseline\left( \left( s_i \right)_{i=0}^t, \baseline\left( \left( s_i \right)_{i=0}^t \right) \right),
			\label{eq:sideEffectAvoidance}
		\end{equation}
		where $t$ is the time step, $s_i$ is the state at time $i$, $\baseline$ is a probability
		distribution of trajectories of the baseline agent, and $\deviationFromBaseline$ is a measure
		of the deviation between a trajectory and a baseline distribution of trajectories.

		In practice $\baseline$ either depends on the initial state and the time step
		(length of trajectory) (Krakovna), or only on the previous state (Turner)	and not both.

		The deviation $\deviationFromBaseline$ is a more involved metric, usually depending on
		relative state	reachability. All approaches up to now only used the last state
		in the trajectories under consideration.
		\label{def:SEA}
	\end{definition}

	Note that the modified reward function is not only a function of the state, but of
	the agent's trajectory and some baseline trajectory distribution.

\section{Offsetting}

	Let us consider two policies. The policy optimal with respect to the base available
	(corrupted) reward $\pi_\corruption$, and the policy optimal with respect to the modified
	reward $\pi_{\modifiedRewardFunction}$.
	We can now define a broad version of offsetting.
	\begin{definition}[Broad offsetting]
		We call the behaviour of an agent following $\pi_{\modifiedRewardFunction}$ that differs
		from the behaviour of an agent following $\pi_\corruption$ \emph{broad offsetting}.
		\label{def:broadOffsetting}
	\end{definition}
	Note that under this definition of offsetting we obviously desire \emph{some} offsetting,
	this was the whole goal of introducing the modified reward in the first place.

	This definition of offsetting also includes interventions in the environment as types
	of offsetting, e.g. in the Sushi example with the baseline being the initial state repeated
	forever and any reasonable deviation measure moving the sushi from the belt is broad offsetting.

	I am under the impression that some researchers understand offsetting differently,
	more based on time -- e.g. Krakovna's formulates it as ``undoing actions''.
	I propose that this definition cannot be made consistent with the examples used
	and some other natural ones.

	\begin{example}[Safety door]
		An agent's task is to turn on a machine in a factory. The factory has a safety door that
		closes whenever the machine is engaged. For some deviation measures the agent will be
		incentivized to block the door from closing if it is impossible to open it afterwards.

		This is clearly not ``undoing'' an action, but calling it offsetting seems appropriate.
	\label{eg:factory}
	\end{example}

	From now I will just write offsetting, omitting the \emph{broad} qualifier.

	\subsection{Desirable vs undesirable offsetting}
		Under the above definition it is clear that some offsetting is desirable, while some is not.
		The formal distinction here is as obvious as useless -- desirable offsetting consists of these
		differences which increase the real reward $\rewardFunction$, while undesirable
		offsetting decreases it. The challenge lies in categorizing these types of offsetting
		without access to the real reward. This requires some assumptions about $\corruption$, as
		well as the baseline.

		\begin{definition}[Goal-correct reward]
			We postulate that the real reward function consists of two components
			\begin{equation*}
				\rewardFunction = \goalRewardFunction + \generalRewardFunction
			\end{equation*}
			where $\goalRewardFunction$ is all the reward received in connection to the goal that
			the agent was supposed to achieve, while $\generalRewardFunction$ is all the background
			reward function implying that vases shouldn't be broken etc.

			% Is this too narrow?
			We call a corrupt reward $\corruption$ \emph{goal-correct} if it is equal
			to $\goalRewardFunction$.
			\label{def:goalCorrect}
		\end{definition}

		Note that this definition is mathematically vacuous, it only works as an intuition,
		and even then it's unclear. Consider Example \ref{eg:factory} -- is the door closing
		a part of the goal in the sense that having a negative or lesser reward if it didn't
		close is required for the reward to be goal-correct?

		\begin{definition}[Human compatible baseline]
			We call a baseline $\baseline$ \emph{human compatible} if the expected cumulative reward
			$\generalRewardFunction$ over it is not lower than the natural evolution's of the environment.
			\label{def:humanCompatible}
		\end{definition}

		I am sneaking in an assumption in the name that lies at the base of the whole
		side-effect avoidance effort -- the environments in which the agent is acting are already
		at least somewhat optimized for humans, so not modifying them is an acceptable action.

		Both the most recent approaches of Krakovna and Turner are trying to pick human compatible
		baselines, usually by making the agent not act. In the case of Krakovna it is not acting
		since the beginning of the trajectory, while in case of Turner it is not acting for a single step.

		Krakovna's position seems to be that if the reward is goal-correct and the baseline is
		human compatible, then all offsetting is desirable. This might be true in theory, but
		due to the unclear boundaries as to what counts as connected to the goal it does not
		really solve the problem in practice. Either we assume that we have to perfectly specify
		\emph{all} intended consequences of the goal being reached (which is probably not feasible
		in sufficiently complicated environments) or some will slip through the cracks.

		The obvious counterargument is that side effect avoidance is only supposed to mitigate
		problems not solve them completely. From this point of view we end up with a \emph{safer}
		AI, even if not with an actually \emph{safe} one.
\section{Onsetting?}
 Do we want to relax the goal-correctness assumptions on the specified reward and try to infer
	a more goal-correct version of the reward?

\end{document}